Working with Daily Climate Model Output Data in R and the \pkg{futureheatwaves} Package

### 2017-05-24 paper_drafts/RJournalRevisions2/response_to_reviewers.txt

Further, we have added a section to the end of the paper, in response to this and other comments
from this reviewer, called "Future directions for working with climate model output in R". In this
section, we have added some text outlining the reviewer's idea that wrappers for these two
speciality languages could be an interesting next step to pursue as R software for working with
climate model output data evolves, including references to existing libraries with wrappers in
other languages (Python and Ruby):

Added text: "One important step in future development of R software to work with climate model output could be the development of R wrappers for some of the existing command line tools available
through the Climate Data Operators (CDO) software (Schulzweida, 2017). Libraries already exist for
Python and Ruby that allow the functionality of CDO tools to be used within these scripting
languages (available from the Max Planck Institute at
https://code.zmaw.de/projects/cdo/wiki/Cdo%7Brbpy%7D). While one R package (`ncdf4.helpers`;
Bronaugh, 2014) already provides R wrappers for a few CDO operators, such functionality could be
extended through future R software to capture more of the full functionality of the CDO toolkit."

Comment: The analysis that this package does is fairly straightforward using `plyr` or other
common R packages / functions. The main function described in this article is the identifying
and summing ranges of events. If a user made it far enough to do the pre-processing of the CMIP5
output data I doubt this would be a computational barrier to further analysis. If authors could
show notable speed or memory requirement improvement then this could be interesting.

Response: As noted in a previous response, the advantage of this package is to improve
convenience for researchers identifying and characterizing sustained, multi-day events in
daily climate model output, especially for analyses that incorporate multiple climate
models and ensemble members. We have added a paragraph to the manuscript making this
motivation for the described package clearer:

### 2017-06-01 paper_drafts/RJournalRevisions2/response_to_reviewers.txt

Thank you very much for the opportunity to revise and resubmit this manuscript. The
reviewers made some excellent suggestions, and we have modified the manuscript to
respond to these. In cases where suggested changes are beyond the scope of the described
package, we have added text to the manuscript to explain how such proposed approaches could
contribute as R programmers continue to evolve software solutions to working with climate model
output in R, and we have tried to make some connections to how R-based approaches to working with
large data continue to evolve in other topical areas with large, array-based data (e.g., flow
cytometry). In addition, we have revised and updated the package that is described by this
manuscript, both in response to reviewer suggestions and also to ensure functionality with recent
updates to `dplyr` and other key imported packages that have undergone major updates since the
original submission. This updated version of the package is available through GitHub
(https://www.github.com/geanders/futureheatwaves), and we will submit this updated package
version to CRAN as soon as we have satisfactorily responded to reviewers.

@@

Response: We appreciate the reviewer's point that the described package does not
resolve all the challenges of working with climate model output data in R.
However, we think that the package described offers an important step forward in making the
ensemble techniques that have become popular in climate science research more accessible to
researchers who are conducting climate impact studies but are not themselves climate scientists.
We have added a "Motivation" subsection to the manuscript section describing the `futureheatwaves`
package to clarify this point:

Added text: "We created the `futureheatwaves` package to aid in identifying, characterizing, and
exploring multi-day extreme events in daily climate model output data. While most of the discrete
tasks involved in identifying and characterizing multi-day extreme events are fairly
straightforward, the full process can be code-intensive, especially for multi-city studies and
studies that test sensitivity to how an event is defined or that incorporate different scenarios of
adaptation in the case of events defined using a threshold relative to community climate. Our aim
in developing this package was therefore to make the full process of identifying and characterizing
these extreme events much more convenient and so facilitate the use of multi-model, multi-ensemble
member analyses in climate impact studies conducted by non-climate scientists."

Further, we have also added a section to the end of the paper ("Future directions for working
with climate model output in R") that describes how future software development in R could
continue to tackle the challenges provided by climate data, to help put the package described in
the current manuscript in context within the evolving use of R for research involving climate
model output data.
